---
title: 'Lab 3: binary logistic regression'
author: "Patrick Pelegri-O'Day"
date: "1/20/2022"
output: html_document
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(palmerpenguins)
library(GGally)
library(broom)
library(jtools)
library(caret)
library(AICcmodavg)
```

### Binary Logistic Regression

#### explore data using `ggpairs`

```{r}
penguins %>% 
  select(species, bill_length_mm:sex) %>% 
  ggpairs(aes(color = species))
```

#### make subset of Chinstrap and Adelie

```{r}
adelie_chinstrap <- penguins %>% 
  filter(species %in% c('Adelie', 'Chinstrap')) %>% 
  mutate(species = fct_drop(species)) %>% 
  select(-island, -year) %>% 
  drop_na()
```

Let's check some trends for variables for these species

```{r}
ggplot(adelie_chinstrap, aes(body_mass_g, flipper_length_mm)) +
  geom_point(aes(color = sex)) +
  facet_wrap(~species)

ggplot(adelie_chinstrap, aes(body_mass_g, bill_length_mm)) +
  geom_point(aes(color = sex)) + 
  facet_wrap(~species)
```

#### Logistic regression in R

```{r}
f1 <- species ~ body_mass_g + flipper_length_mm + sex

ad_chin_blr1 <- glm(formula = f1,
                    data = adelie_chinstrap,
                    family = 'binomial')
```

look at the results:
```{r}
ad_chin_blr1

summary(ad_chin_blr1)

blr1_tidy <- tidy(ad_chin_blr1)
```

```{r}
ggplot(adelie_chinstrap, aes(species, body_mass_g)) +
  geom_jitter(aes(color = sex))

ggplot(adelie_chinstrap, aes(species, flipper_length_mm)) +
  geom_jitter(aes(color = sex))
```

```{r}
blr1_fitted <- ad_chin_blr1 %>% 
  broom::augment(type.predict = 'response')
```

```{r}
ggplot(data = blr1_fitted, aes(flipper_length_mm, .fitted)) +
  geom_point(aes(color = species, shape = sex)) +
  geom_smooth(se = FALSE) +
  labs(x = 'Flipper length (mm)',
       y = 'Probability of outcome "Chinstrap"')
```

Visualize model outcomes using `jtools::effect_plot()` function

```{r}
effect_plot(ad_chin_blr1,
            pred= flipper_length_mm,
            interval = TRUE,
            y.label = "probability of chinstrap")

effect_plot(ad_chin_blr1,
            pred= body_mass_g,
            interval = TRUE,
            y.label = "probability of chinstrap")
```

#### Predict species based on new values

What is the probability that a female penguin with 3410g body mass and flipper length of 192mm will be "chinstrap"?
```{r}
ex1 <- predict(ad_chin_blr1,
               data.frame(sex = 'female',
                          body_mass_g = 3410,
                          flipper_length_mm = 192),
               type = 'response') # 'response' converts from log odds to probability
```

```{r}
new_df <- data.frame(
  sex = c('male', 'male', 'female'),
  body_mass_g = c(3298, 4100, 3600),
  flipper_length_mm = c(212, 175, 180)
)

ex2 <- predict(ad_chin_blr1,
               new_df,
               type = 'response')
```

#### New model - bill length

```{r}
f2 <- species ~ bill_length_mm + body_mass_g

ad_chin_blr2 <- glm(formula = f2,
                    data = adelie_chinstrap,
                    family = "binomial")
```

```{r}
ad_chin_blr2

summary(ad_chin_blr2)
```

```{r}
ggplot(adelie_chinstrap, aes(bill_length_mm, body_mass_g)) +
  geom_point(aes(color = species))
```

```{r}
effect_plot(ad_chin_blr2,
            pred = bill_length_mm,
            interval = TRUE,
            y.label = "probability of 'Chinstrap'")

effect_plot(ad_chin_blr2,
            pred = body_mass_g,
            interval = TRUE,
            y.label = "probability of 'Chinstrap'")
```

### Model selection

First look at AIC values

```{r}
AICcmodavg::aictab(list(ad_chin_blr1, ad_chin_blr2))
```

Let's do 10-fold cross-validation, use prediction accuracy as our metric.

```{r}
set.seed(123)

n_folds <- 10
folds <- rep(1:n_folds, length.out = nrow(adelie_chinstrap))

ad_chin_kfold  <- adelie_chinstrap %>% 
  mutate(fold = sample(folds, size = n(), replace = FALSE))
```

```{r}
pred_acc <- function(x, y) {
  accurate <- ifelse(x == y, 1, 0)
  return(mean(accurate, na.rm = TRUE))
}
```

```{r}
results_df <- data.frame()

for(i in 1:n_folds) {
  kfold_test <- ad_chin_kfold %>% 
    filter(fold == i)
  kfold_train <- ad_chin_kfold %>% 
    filter(fold != i)
  
  kfold_blr1 <- glm(f1, data = kfold_train, family = 'binomial')
  kfold_blr2 <- glm(f2, data = kfold_train, family = 'binomial')
  
  kfold_pred <- kfold_test %>% 
    mutate(blr1 = predict(kfold_blr1, kfold_test, type = 'response'),
           blr2 = predict(kfold_blr2, ., type = 'response')) %>% 
  mutate(pred1 = ifelse(blr1 > 0.5, 'Chinstrap', 'Adelie'),
         pred2 = ifelse(blr2 > 0.5, 'Chinstrap', 'Adelie'))
  
  kfold_accuracy <- kfold_pred %>% 
    summarize(blr1_acc = pred_acc(species, pred1),
              blr2_acc = pred_acc(species, pred2))
  
  results_df <- bind_rows(results_df, kfold_accuracy)
}

results_df %>% 
  summarize(blr1_acc = mean(blr1_acc),
            blr2_acc = mean(blr2_acc))
```

Use the `caret` package to automate the kfold cross validation

```{r}
set.seed(123)

tr_ctrl <- trainControl(method = 'repeatedcv', number = 10, repeats = 10)

### train the model
model1 <- train(f1, data = adelie_chinstrap,
                method = 'glm', family = 'binomial',
                trControl = tr_ctrl)

model1

model2 <- train(f2, data = adelie_chinstrap,
                method = 'glm', family = 'binomial',
                trControl = tr_ctrl)

model2
```




